---
title: "Hands-on Exercise 2: Chloropeth"
execute: 
  warning: false
  #eval: false
  
date: "r Sys.Date()`"
---

Chloropeth Map (Heatmap)

A choropleth map is a type of thematic map in which areas are shaded or patterned in proportion to a statistical variable that represents an aggregate summary of a geographic characteristic within each area, such as population or per-capita income.

### Classified choropleth map

-   Choropleth maps can be either classified or unclassified.

-   A classed choropleth map combines areal units into a smaller number of groups. Interval levels may vary, but typically 4 to 7 are used in a map. There is different classification techniques used to divide up the intervals.

### Unclassified choropleth map

-   Unclassed choropleth maps are similar to classed choropleth maps; however, unclassed choropleth maps do not have an averaged statistic towards each particular colour.

### Choosing an appropriate number of classes

-   Sturges formula

-   Interval not more than 4, not more than 12 colours, 7-8 shade same color

Data classification

-   Equal int(Range of bin same), Quantile(Divide into 20%'s, easy to see density), Natural breaks(Jenks)- make sure data is not highly skewed, manual - map the standard deviation - useful if data represents normal distribution

Colour scheme

-   ColorBrewer - systematic color scheme

-   Color split int (Nominal, Sequential - When all positive or all negative, Diverging - use when some positive some negative)

The code chunk below will be used to install and load these packages in RStudio.

```{r}
pacman::p_load(sf, tmap, tidyverse)

```

Importing Geospatial Data into R
```{r}
mpsz <- st_read(dsn = "../data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

Importing Attribute Data into R
```{r}
popdata <- read_csv("../data/aspatial/respopagesextod2011to2020.csv")
```

Data Preparation
```{r}
column_names <- names(popdata)

# Print all column names
cat("Column Names: ", paste(column_names, collapse = ", "), "\n")

```

Checking what does pivot_wider do
```{r}
print(popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) )

```

 Data wrangling
 The following data wrangling and transformation functions will be used:

pivot_wider() of tidyr package, and
mutate(), filter(), group_by() and select() of dplyr package
```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

Joining the attribute data and geospatial data

  
```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = list(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

Next, left_join() of dplyr is used to join the geographical data and attribute table using planning subzone name e.g. SUBZONE_N and SZ as the common identifier.

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

left_join() of dplyr package is used with mpsz simple feature data frame as the left data table is to ensure that the output will be a simple features data frame.
```{r}
write_rds(mpsz_pop2020, "../data/rds/mpszpop2020.rds")
```

Choropleth Mapping Geospatial Data Using tmap
Choropleth mapping involves the symbolisation of enumeration units, such as countries, provinces, states, counties or census units, using area patterns or graduated colors. For example, a social scientist may need to use a choropleth map to portray the spatial distribution of aged population of Singapore by Master Plan 2014 Subzone Boundary.

Two approaches can be used to prepare thematic map using tmap, they are:

Plotting a thematic map quickly by using qtm().
Plotting highly customisable thematic map by using tmap elements.

2.4.1 Plotting a choropleth map quickly by using qtm()
```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
print(mpsz_pop2020)
```

2.4.2 Creating a choropleth map by using tmap’s elements
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))

```


2.4.2.1 Drawing a base map
```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

2.4.2.2 Drawing a choropleth map using tm_polygons()
```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

2.4.2.3 Drawing a choropleth map using tm_fill() and *tm_border()**
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```


2.4.3 Data classification methods of tmap

Jenks (Quantile??)
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```

Equal
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```

2.4.3.2 Plotting choropleth map with custome break
```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

Custom break
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(alpha = 0.5)
```

2.4.4 Colour Scheme
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)

```

Reverse colour
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)

```


2.4.5 Map Layouts
2.4.5.1 Map Legend
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)


```


2.4.5.2 Map style
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

2.4.5.3 Cartographic Furniture - includes compass, scales etc
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

2.4.6 Drawing Small Multiple Choropleth Maps
```{r}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

Differentaesthetic argument
```{r}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

2.4.6.2 By defining a group-by variable in tm_facets()
```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

2.4.6.3 By creating multiple stand-alone maps with tmap_arrange()
```{r}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

2.4.7 Mappping Spatial Object Meeting a Selection Criterion
```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```