---
title: "Hands on assignment 11 - Glenn"
execute: 
  warning: false
  #eval: false
  
date: "r Sys.Date()`"
---
# Preparing the data
```{r}
pacman::p_load(tmap, sf, DT, stplanr, tidyverse)
```

```{r}
odbus <- read_csv("data/aspatial/origin_destination_bus_202210.csv")
```

```{r}
odbus$ORIGIN_PT_CODE <- as.factor(odbus$ORIGIN_PT_CODE)
odbus$DESTINATION_PT_CODE <- as.factor(odbus$DESTINATION_PT_CODE) 
```

## Extracting the study data
For the purpose of this exercise, we will extract commuting flows on weekday and between 6 and 9 o’clock.
```{r}
odbus6_9 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 6 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE,
           DESTINATION_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
write_rds(odbus6_9, "data/rds/odbus6_9.rds")
```

```{r}
odbus6_9 <- read_rds("data/rds/odbus6_9.rds")
```

# Working with Geospatial Data
## Importing
```{r}
busstop <- st_read(dsn = "data/geospatial",
                   layer = "BusStop") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz <- st_read(dsn = "data/geospatial",
                   layer = "MPSZ-2019") %>%
  st_transform(crs = 3414)
```

```{r}
mpsz <- write_rds(mpsz, "data/rds/mpsz.rds")
```

# Geospatial data wrangling
## Combining Busstop and mpsz
Code chunk below populates the planning subzone code (i.e. SUBZONE_C) of mpsz sf data frame into busstop sf data frame.
```{r}
busstop_mpsz <- st_intersection(busstop, mpsz) %>%
  select(BUS_STOP_N, SUBZONE_C) %>%
  st_drop_geometry()
```
```{r}
write_rds(busstop_mpsz, "data/rds/busstop_mpsz.rds")  
```

```{r}
od_data <- left_join(odbus6_9 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C,
         DESTIN_BS = DESTINATION_PT_CODE)
```

Before continue, it is a good practice for us to check for duplicating records.
```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

If duplicated records are found, the code chunk below will be used to retain the unique records.
```{r}
od_data <- unique(od_data)
```

Next, we will update od_data data frame cwith the planning subzone codes.
```{r}
od_data <- left_join(od_data , busstop_mpsz,
            by = c("DESTIN_BS" = "BUS_STOP_N")) 

duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()

od_data <- unique(od_data)

od_data <- od_data %>%
  rename(DESTIN_SZ = SUBZONE_C) %>%
  drop_na() %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>%
  summarise(MORNING_PEAK = sum(TRIPS))
```

It is time to save the output into an rds file format.
```{r}
write_rds(od_data, "data/rds/od_data_fii.rds")
```

```{r}
od_data_fii <- read_rds("data/rds/od_data_fii.rds")
```

# Visualising Spatial Interaction
## Removing intra-zonal flows
```{r}
od_data_fij <- od_data[od_data$ORIGIN_SZ!=od_data$DESTIN_SZ,]
```

```{r}
write_rds(od_data_fij, "data/rds/od_data_fij.rds")
```

```{r}
od_data_fij <- read_rds("data/rds/od_data_fij.rds")
```

## Creating desire lines
```{r}
flowLine <- od2line(flow = od_data_fij, 
                    zones = mpsz,
                    zone_code = "SUBZONE_C")
```
```{r}
write_rds(flowLine, "data/rds/flowLine.rds")
```
```{r}
flowLine <- read_rds("data/rds/flowLine.rds")
```


## Visualising the desire lines
```{r}
tm_shape(mpsz) +
  tm_polygons() +
flowLine %>%  
tm_shape() +
  tm_lines(lwd = "MORNING_PEAK",
           style = "quantile",
           scale = c(0.1, 1, 3, 5, 7, 10),
           n = 6,
           alpha = 0.3)
```







# Calibrating Spatial Interaction Models with R
# Getting Started
```{r}
pacman::p_load(tmap, sf, sp,
               performance, reshape2,
               ggpubr, tidyverse)
```

# Computing Distance Matrix
```{r}
mpsz <-st_read(dsn = "data/geospatial", 
                layer = "MPSZ-2019")

```

## Converting from sf data.table to SpatialPolygonsDataFrame
```{r}
mpsz_sp <- as(mpsz, "Spatial")
mpsz_sp
```

## Computing the distance matrix
```{r}

dist <- spDists(mpsz_sp, 
                longlat = FALSE)
```

##  Labelling column and row heanders of a distance matrix
```{r}
sz_names <- mpsz$SUBZONE_C
```

```{r}
colnames(dist) <- paste0(sz_names)
rownames(dist) <- paste0(sz_names)
```


## Pivoting distance value by SUBZONE_C
```{r}
distPair <- melt(dist) %>%
  rename(dist = value)
head(distPair, 10)
```


## Updating intra-zonal distances
In this section, we are going to append a constant value to replace the intra-zonal distance of 0.

First, we will select and find out the minimum value of the distance by using summary().
```{r}
distPair %>%
  filter(dist > 0) %>%
  summary()

```

Next, a constant distance value of 50m is added into intra-zones distance.
```{r}
distPair$dist <- ifelse(distPair$dist == 0,
                        50, distPair$dist)
```

The code chunk below is used to rename the origin and destination fields.
```{r}
distPair <- distPair %>%
  rename(orig = Var1,
         dest = Var2)
```

```{r}
write_rds(distPair, "data/rds/distPair.rds") 
```

```{r}
distPair <- read_rds("data/rds/distPair.rds")
```

# Preparing flow data
```{r}
od_data_fii <- read_rds("data/rds/od_data_fii.rds")
```

Next, we will compute the total passenger trip between and within planning subzones by using the code chunk below. The output is all flow_data.
```{r}
flow_data <- od_data_fii %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>% 
  summarize(TRIPS = sum(MORNING_PEAK)) 
```

##  Separating intra-flow from passenger volume df
Code chunk below is used to add three new fields in flow_data dataframe.
```{r}
flow_data$FlowNoIntra <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0, flow_data$TRIPS)
flow_data$offset <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0.000001, 1)
```

## Combining passenger volume data with distance value
Before we can join flow_data and distPair, we need to convert data value type of ORIGIN_SZ and DESTIN_SZ fields of flow_data dataframe into factor data type.
```{r}
flow_data$ORIGIN_SZ <- as.factor(flow_data$ORIGIN_SZ)
flow_data$DESTIN_SZ <- as.factor(flow_data$DESTIN_SZ)
```

Now, left_join() of dplyr will be used to flow_data dataframe and distPair dataframe. The output is called flow_data1.
```{r}
flow_data1 <- flow_data %>%
  left_join (distPair,
             by = c("ORIGIN_SZ" = "orig",
                    "DESTIN_SZ" = "dest"))
```

# Preparing Origin and Destination Attributes
## Importing population data
```{r}
pop <- read_csv("data/aspatial/pop.csv")
```

## Geospatial data wrangling
```{r}
pop <- pop %>%
  left_join(mpsz,
            by = c("PA" = "PLN_AREA_N",
                   "SZ" = "SUBZONE_N")) %>%
  select(1:6) %>%
  rename(SZ_NAME = SZ,
         SZ = SUBZONE_C)
```

## Preparing origin attribute
```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(ORIGIN_SZ = "SZ")) %>%
  rename(ORIGIN_AGE7_12 = AGE7_12,
         ORIGIN_AGE13_24 = AGE13_24,
         ORIGIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

## Preparing destination attribute
```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(DESTIN_SZ = "SZ")) %>%
  rename(DESTIN_AGE7_12 = AGE7_12,
         DESTIN_AGE13_24 = AGE13_24,
         DESTIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

We will called the output data file SIM_data. it is in rds data file format.
```{r}
write_rds(flow_data1, "data/rds/flow_data_6-9.rds")
```

# Calibrating Spatial Interaction Models
## Importing the modelling data
```{r}
SIM_data <- read_rds("data/rds/flow_data_6-9.rds")
```

## Visualising the dependent variable
```{r}
ggplot(data = SIM_data,
       aes(x = TRIPS)) +
  geom_histogram()
```

Next, let us visualise the relation between the dependent variable and one of the key independent variable in Spatial Interaction Model, namely distance.
```{r}
ggplot(data = SIM_data,
       aes(x = dist,
           y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)
```
On the other hand, if we plot the scatter plot by using the log transformed version of both variables, we can see that their relationship is more resemble linear relationship.
```{r}
ggplot(data = SIM_data,
       aes(x = log(dist),
           y = log(TRIPS))) +
  geom_point() +
  geom_smooth(method = lm)
```

## Checking for variables with zero values
The print report above reveals that variables ORIGIN_AGE7_12, ORIGIN_AGE13_24, ORIGIN_AGE25_64,DESTIN_AGE7_12, DESTIN_AGE13_24, DESTIN_AGE25_64 consist of 0 values.

In view of this, code chunk below will be used to replace zero values to 0.99.
```{r}
SIM_data$DESTIN_AGE7_12 <- ifelse(
  SIM_data$DESTIN_AGE7_12 == 0,
  0.99, SIM_data$DESTIN_AGE7_12)
SIM_data$DESTIN_AGE13_24 <- ifelse(
  SIM_data$DESTIN_AGE13_24 == 0,
  0.99, SIM_data$DESTIN_AGE13_24)
SIM_data$DESTIN_AGE25_64 <- ifelse(
  SIM_data$DESTIN_AGE25_64 == 0,
  0.99, SIM_data$DESTIN_AGE25_64)
SIM_data$ORIGIN_AGE7_12 <- ifelse(
  SIM_data$ORIGIN_AGE7_12 == 0,
  0.99, SIM_data$ORIGIN_AGE7_12)
SIM_data$ORIGIN_AGE13_24 <- ifelse(
  SIM_data$ORIGIN_AGE13_24 == 0,
  0.99, SIM_data$ORIGIN_AGE13_24)
SIM_data$ORIGIN_AGE25_64 <- ifelse(
  SIM_data$ORIGIN_AGE25_64 == 0,
  0.99, SIM_data$ORIGIN_AGE25_64)
```

## Unconstrained Spatial Interaction Model
In this section, you will learn how to calibrate an unconstrained spatial interaction model by using glm() of Base Stats. The explanatory variables are origin population by different age cohort, destination population by different age cohort (i.e. ORIGIN_AGE25_64) and distance between origin and destination in km (i.e. dist).
```{r}
uncSIM <- glm(formula = TRIPS ~ 
                log(ORIGIN_AGE25_64) + 
                log(DESTIN_AGE25_64) +
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
uncSIM
```

##  R-squared function
```{r}
CalcRSquared <- function(observed,estimated){
  r <- cor(observed,estimated)
  R2 <- r^2
  R2
}
```

```{r}
CalcRSquared(uncSIM$data$TRIPS, uncSIM$fitted.values)
```

```{r}
r2_mcfadden(uncSIM)
```

## Origin (Production) constrained SIM
```{r}
orcSIM <- glm(formula = TRIPS ~ 
                 ORIGIN_SZ +
                 log(DESTIN_AGE25_64) +
                 log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(orcSIM)
```

```{r}
CalcRSquared(orcSIM$data$TRIPS, orcSIM$fitted.values)
```

## Destination constrained
```{r}
decSIM <- glm(formula = TRIPS ~ 
                DESTIN_SZ + 
                log(ORIGIN_AGE25_64) + 
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(decSIM)
```


```{r}
CalcRSquared(decSIM$data$TRIPS, decSIM$fitted.values)
```

## Doubly constrained
```{r}
dbcSIM <- glm(formula = TRIPS ~ 
                ORIGIN_SZ + 
                DESTIN_SZ + 
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(dbcSIM)
```

```{r}
CalcRSquared(dbcSIM$data$TRIPS, dbcSIM$fitted.values)
```

## Model comparison
In this sub-section, you will learn how to use compare_performance() of performance package
First of all, let us create a list called model_list by using the code chun below.
```{r}
model_list <- list(unconstrained=uncSIM,
                   originConstrained=orcSIM,
                   destinationConstrained=decSIM,
                   doublyConstrained=dbcSIM)
```

```{r}
compare_performance(model_list,
                    metrics = "RMSE")
```

SMALLEST VALUE = best model

## Visualising fitted values
In this section, you will learn how to visualise the observed values and the fitted values.
```{r}
df <- as.data.frame(uncSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(uncTRIPS = "uncSIM$fitted.values")
```

```{r}
df <- as.data.frame(orcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(orcTRIPS = "orcSIM$fitted.values")
```

Repeat the same step by for Doubly Constrained SIM (i.e. dbcSIM)
```{r}
df <- as.data.frame(dbcSIM$fitted.values) %>%
  round(digits = 0)

SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(dbcTRIPS = "dbcSIM$fitted.values")

unc_p <- ggplot(data = SIM_data,
                aes(x = uncTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

orc_p <- ggplot(data = SIM_data,
                aes(x = orcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dec_p <- ggplot(data = SIM_data,
                aes(x = decTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dbc_p <- ggplot(data = SIM_data,
                aes(x = dbcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)
```

```{r}
ggarrange(unc_p, orc_p, dbc_p,
          ncol = 2,
          nrow = 2)
```

